---
title: "In Class Exercises - Sorting and Filtering"
author: "Jenna Baer"
date: "`r format(Sys.time(), '%B %d, %Y')`"
format:
  html:
    self-contained: true
    toc: true
    toc-location: left
---

## Setup and Loading Data

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

# Load the packages we'll need
# Remember, think of this like packing for our trip
# (We don't take the whole closet, only what we need)
library(tidyverse)
library(lubridate)
```

Run this line below load the data for this assignment. It will load a dataframe called "impeach," containing U.S. House Democrats and whether they supported the first impeachment of Pres. Trump, along with when they announced such support.

```{r}

impeach <- read_csv("https://docs.google.com/spreadsheets/d/e/2PACX-1vRh8d5JaDqBtByzNw2JZF3idaACobDhhk-p7chJoktA0HawELHFOvjQOqCpzGA4MGONvPlR7GASqW-K/pub?gid=1765341510&single=true&output=csv")


```

Let's see what we have:

```{r}

impeach


```

## Exercise Questions

*For the questions below, write your R code to return the requested records in the code chunk provided.*

1)  The column "for_impeachment" indicates whether the member has publicly called for an impeachment inquiry. Filter to return only the ones where the answer is NO. 
I remembered from the in class activity to use the filter function and apply it to the whole data set. After setting that up I filtered the for impeachment column because that is the variable of interest here.

```{r}

filter(impeach, for_impeachment == "NO")






```

2)  Filter to return only results where a member is both against impeachment, and comes from a district that President Trump won in 2016 (which is noted in the "p16winningparty" column). 

I decided to use the pipe function here because we're combining variables through the filter function. I set up the variable for Trump winning a district in 2016 to be equivalent to "R" because that meant a Republican won the district.
```{r}

impeach %>%
  filter(for_impeachment == "NO", p16winningparty == "R")



```

3)  Filter for only results where a member is against impeachment, comes from a district that President Trump won in 2016 (which is noted in the "p16winningparty" column), and also comes from a district that Mitt Romney won in 2012 ("p12winningparty").

I just added onto the pipe function using the same format as the previous question.

```{r}

impeach %>%
  filter(for_impeachment == "NO", p16winningparty == "R", p12winningparty == "R")



```

4)  Filter for only results from September 2019 where a member is a YES for impeachment.

I again knew to use the pipe and filter functions because we're narrowing the variables were interested in. I also figured it made the most sense to use the date_year and date_month variables independently because you didn't ask for the entire date/ specific day.
```{r}

impeach %>%
  filter(for_impeachment == "YES", date_year == 2019, date_month == 9)



```

5)  Filter for only results where a member is a YES for impeachment and is from a district where Clinton won more than 70 percent of the vote in 2016 (found in column "clinton_percent").

```{r}





```

6)  Sort the entire dataframe based on the percentage of a district that has a bachelor's degree or higher ("pct_bachelors"), from lowest to highest.

```{r}






```

7)  Sort the just those who are NO on impeachment based on the percentage of a district that has a bachelor's degree or higher ("pct_bachelors"), from lowest to highest.

```{r}





```

8)  Sort the just those who are NO on impeachment based on the percentage of a district that has a bachelor's degree or higher ("pct_bachelors"), from lowest to highest.

Then filter those records by only those whose bachelor's percentage is below the national average (found in the "pct_bachelors_compared_to_national" column).

```{r}






```

9)  Filter for only members from New Jersey who are NO on impeachment.

```{r}





```

10) Filter for those who were YES on impeachment, with a declared date prior to 2019. So only those with dates before 2019. Then sort those so that the highest Clinton vote percentages are at the top.

```{r}





```
